---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
---


##Loading and preprocessing the data

```{r, message=FALSE}
library(dplyr)
library(lubridate)
library(lattice)

activity <- read.csv("./RepData_PeerAssessment1-master/activity.csv")
activity$date <- as.Date(activity$date)
```
## What is mean total number of steps taken per day?

Calculate the total number of steps taken per day

```{r}
day <- 
  activity %>%
  group_by(date) %>%
  summarize(totalsteps = sum (steps))
```

Make a histogram of the total number of steps taken each day

```{r}
with(day, hist(totalsteps, col = "green", breaks = 10, main = "Histogram of total steps", xlab = "Total steps"))
rug(day$totalsteps)
```

Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean(day$totalsteps, na.rm = T)
median(day$totalsteps, na.rm = T)
```

## What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
average <-
  activity %>%
  group_by(interval) %>%
  summarize(average = mean(steps, na.rm = T))
```

```{r}
with(average, plot(interval, average, type = "l"))
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
row_max <- filter(average, average == max(average))
row_max$interval
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset 

```{r}
row_NA <- is.na(activity$steps)
length(which(row_NA==TRUE))
```

Filling in all of the missing values.
The missing values are replaced by the mean values calculated before

```{r}
na_activity <- filter(activity, is.na(steps))
new_na_activity <-
  average %>%
  merge(na_activity, by = "interval") %>%
  select(average, date, interval) %>%
  rename(steps = average)
```

Create a new dataset that is equal to the original dataset but with the missing data filled in

```{r}
new_activity <- bind_rows(filter(activity, !is.na(steps)), new_na_activity)
new_activity <- arrange(new_activity, date, interval)
```

Calculate histogram

```{r}
new_day <- 
  new_activity %>%
  group_by(date) %>%
  summarize(total = sum (steps))
```
```{r}
hist(new_day$total, col = "green", breaks = 10, main = "Histogram of total steps", xlab = "Total steps")
rug(new_day$total)
```

Calculate mean and median

```{r}
mean(new_day$total)
median(new_day$total)
```
#### Do these values differ from the estimates from the first part of the assignment?

The difference is minimal to calculate the mean and the median due to the method used to replace the missing values

#### What is the impact of imputing missing data on the estimates of the total daily number of steps?

There are more days with a maximum number of steps. When replacing the unknown data we have added steps to the total

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day

```{r}
weekend <- 
  new_activity %>%
  mutate(day = wday(date, label = TRUE), week = "Weekend") %>%
  filter(day == "Sat" | day == "Sun") %>%
  group_by(interval, week) %>%
  summarize(average = mean(steps)) 

weekday <- 
  new_activity %>%
  mutate(day = wday(date, label = TRUE), week = "Weekday") %>%
  filter(day != "Sat" , day != "Sun") %>%
  group_by(interval, week) %>%
  summarize(average = mean(steps))

  
week <- bind_rows(weekend, weekday)
```

```{r}
xyplot(average ~ interval | week, data = week, layout = c(1, 2), type = "l")
```

